%% Deletion-Completeness for LLM Agent Memory — AAAI 2027 submission (DRAFT)
%% Drafting order (web-brain locked): §5 Results FIRST. Other sections are stubs.
%% Requires the AAAI author kit style file (aaai2027.sty) on the TeX path.
%%
%% ── TITLE OPTIONS (web brain to pick; all name the decomposition + planner,
%%    none use "benchmark/audit/evaluation" — ForgetEval owns those words) ──
%% (T1) Decomposing Recoverability: Minimal Co-Deletion for the
%%      Right to be Forgotten in LLM Agent Memory
%% (T2) Why Deleted Facts Recur: A Causal Recoverability Decomposition
%%      and Co-Deletion Planner for Agent Memory
%% (T3) Certifying Deletion-Completeness in LLM Agent Memory: Recoverability
%%      Decomposition, Minimal Co-Deletion, and the World Recall

\documentclass[letterpaper]{article}
% TODO(AAAI kit): submission vs camera-ready. This uses \usepackage{aaai2027}
%   (camera-ready / non-anonymous mode). For AAAI double-blind REVIEW submission,
%   change to \usepackage[submission]{aaai2027}, anonymize the \author and
%   \affiliations block, and remove the /Author line from \pdfinfo. See AAAI_SETUP.md.
\usepackage[submission]{aaai2027}  % AAAI author kit (anonymous submission mode for double-blind review)
% NOTE(AAAI kit): aaai2027.sty auto-loads all required fonts (newtxtext for serif,
%   helvet for sans-serif, courier for monospace). The AAAI-27 kit forbids loading
%   \usepackage{times}, \usepackage{helvet}, \usepackage{courier}, or any other font
%   package (times in particular conflicts with newtxtext), so those three lines were
%   removed here for compliance. See AAAI_SETUP.md.
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}   % \varnothing, AMS symbols
\usepackage{dsfont}    % \mathds{1} indicator (amssymb's \mathbb has no digit; AAAI forbids bbm)
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{natbib}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}

% NOTE(AAAI kit): the AAAI-27 kit's sample \pdfinfo contains ONLY /TemplateVersion.
%   /Title and /Author are optional metadata; for an anonymous [submission] build,
%   REMOVE the /Author line below to preserve double-blind review. See AAAI_SETUP.md.
\pdfinfo{
/TemplateVersion (2027.1)
}

\setcounter{secnumdepth}{2}

\title{\textsc{Salman}: Stored-memory Auditing of Leakage, Minimal Co-Deletion,\\
       and Nondeletability for the Right to Be Forgotten in LLM Agent Memory}
\author{
    Anonymous Submission
}
\affiliations{
}

\begin{document}
\maketitle

\begin{abstract}
Ask a paging LLM agent to forget a fact and it may scrub the working-memory block
it reasons over while silently leaving the fact in its archival store: the deletion
is acknowledged, looks complete, and is not. More generally, deleting a fact from
an LLM agent's memory does not mean the agent has forgotten it. Existing audits
check whether a forget command removes a value from retrieval, which is
\emph{necessary but not sufficient}, because the value can persist in a derived
artifact or be re-derived from facts that were never deleted. We formalize
\emph{deletion-completeness} for agent memory, decomposing post-deletion
recoverability into three causally distinct channels (residual survival,
re-derivation from surviving facts, and world recall, the value recoverable from the
base model alone) and give a planner that computes the minimal co-deletion
closing the deletable channels, emitting an auditable certificate. Because the
world recall cannot be deleted, certification has a hard limit: under the
strongest modeled adversary, \textbf{86 of 250} facts cannot be certified erased
(at threshold $\tau{=}0.1$), even with zero residual survival and no helpful
co-deletion. The same stack, run across three memory architectures (a
deduplication pipeline, a bi-temporal knowledge graph, and a paging agent), finds
residual survival in all three through different by-design mechanisms: the gap is a
property of how agent memory derives and retains facts, not of any single system.
\end{abstract}

%========================================================================
\section{Introduction}\label{sec:intro}

A user invokes their right to be forgotten \citep{gdpr2016} and asks an LLM agent
to delete their home address. A modern agent with long-term memory (a paging architecture such as
MemGPT/Letta) acknowledges and rewrites the working-memory block it is currently
reasoning over. But the address was also written to the agent's archival vector
store in an earlier session; reasoning only over its working context, the agent
never touches it, and weeks later an ordinary retrieval surfaces the address again.
The deletion was confirmed, looked complete, and was not.

This is not a one-off bug. Even a \emph{faithful} delete of a record need not erase
the fact it carried: the fact can survive in a derived artifact the system built
from it (a duplicate, a summary, a consolidated note) or be re-derived from facts
that remain, such as a salary from a job title plus a public pay band, or a
diagnosis from stored symptoms. The most thorough current audits of agent
forgetting \citep{forgeteval} check whether a forget command removes a value from
top-$k$ retrieval. That check is \emph{necessary but not sufficient}: it does not
ask \emph{why} a deleted fact returns, nor \emph{what minimal further deletion}
would actually close the leak. Those two questions are the subject of this
paper.\footnote{The authors used generative-AI assistants to help draft and edit prose
and code. All definitions, proofs, experiments, reported numbers, and citations were
verified by the authors, who take full responsibility for the content, consistent with
AAAI's policy on the use of generative AI.}

We cast deletion-completeness as whether a deleted fact remains recoverable, and
decompose recoverability into three causally distinct channels: residual survival
(a surviving artifact still contains the value), re-derivation (surviving facts
entail it), and a world recall (the base model recovers it from world knowledge
alone). The first two are closable by deleting more; the third is not. A greedy
planner computes the minimal additional co-deletion that closes the closable
channels, and a certificate records what deletion achieved, at what collateral
cost, and when the world recall makes erasure impossible to
certify. Concretely, we contribute:

\begin{itemize}
\item a formal, \emph{adversary-relative} deletion-completeness criterion for agent
memory, strictly stronger than dependency-bounded database erasure, decomposing
recoverability into residual, re-derivation, and world-recall channels (\S\ref{sec:formulation});
\item a minimal co-deletion planner for the resulting NP-hard problem, reaching
completeness at mean collateral $k=1.03$ with zero spurious deletions, plus an
auditable certificate that separates what deletion achieved from what it cannot
(\S\ref{sec:method},\,\S\ref{sec:res-planner});
\item a \emph{measured} world recall and its limit result: under the strongest
modeled adversary, $86/250$ facts cannot be certified erased even at zero residual
survival. Measured as a recovery rate over $8$ samples per fact,
the count is a Monte-Carlo estimate stable to $\pm1$--$2$ facts at the $n{=}8$-sample boundary
(\S\ref{sec:res-rho});
\item evidence of generality: the same probe/planner/certificate stack, run on
three memory architectures, exposes residual survival in all three through
different by-design mechanisms, including the agent-loop failure above that prior
audits bypass (\S\ref{sec:res-converge}).
\end{itemize}

We call the resulting framework \textsc{Salman}: \textbf{S}tored-memory
\textbf{A}uditing of \textbf{L}eakage, \textbf{M}inimal co-deletion, \textbf{a}nd
\textbf{N}ondeletability.

The breadth across systems is evidence for a stronger, causally-decomposed notion
of erasure. This is not a claim to be first to evaluate these systems, since ForgetEval
\citep{forgeteval} already benchmarks all three, but a stronger question asked of
them. What is new is asking, and answering, why deleted facts return and what it
costs to make them stay gone.

\section{Related Work}\label{sec:related}

\paragraph{Auditing forgetting in agent memory.} Agent memory is now routinely
benchmarked for long-term \emph{retention}, whether an assistant recalls facts
across long histories \citep{maharana2024locomo,wu2025longmemeval}; its inverse,
whether a fact can be reliably \emph{forgotten}, is far less examined. The closest
prior work is \textsc{ForgetEval} \citep{forgeteval}, the most thorough existing
audit of forgetting in deployed agent memory: a careful retrieval-based protocol
spanning Mem0, Graphiti, Letta, and roughly a dozen further systems. We build directly
on it. Its retrieval-based criterion is the \emph{necessary-but-not-sufficient}
gap of \S\ref{sec:intro}. Our
contribution is therefore the stronger, channel-decomposed criterion, not the
system coverage ForgetEval already establishes; correspondingly, our Mem0 figures
(recoverability by \emph{any} channel) are complementary to its top-$k$
forget-command rates, not in tension with them. Two of our results extend it
directly. ForgetEval deliberately bypasses Letta's agent loop (addressing
archival endpoints to keep the model out of the recall path), which is exactly the
loop we exercise and find surface-incomplete (\S\ref{sec:res-converge}); and its
own per-system observations (Graphiti shedding synthesized fact strings, Mem0's
router favoring link-and-keep) name the mechanisms our decomposition isolates,
independently corroborating the convergence result. A complementary governance
direction attaches \emph{provenance} to agent memory: \textsc{MemLineage}
\citep{memlineage} signs each entry and tracks its derivation lineage so that actions
descending from untrusted content can be gated. It shares our premise that memories
derive from one another, but it enforces \emph{trust} while keeping the content in
place, whereas we \emph{remove} a fact and certify the removal.

\paragraph{Unlearning methods.} Classical machine unlearning removes a training example
by retraining or editing model weights
\citep{bourtoule2021unlearning,cao2015towards,nguyen2022survey}; recent work shows that
such isolated-fact removal overlooks \emph{latent inferential dependencies}, so a
``forgotten'' fact persists through correlated surviving knowledge \citep{dollmsforget},
which is the re-derivation channel we isolate and then close. Closest to us,
\textsc{Agentic Unlearning} \citep{agenticunlearning} also splits parametric from memory
recovery (its ``backflow'') and, on the memory side, prunes dependent artifacts by a
dependency closure that needs no weight access. We differ by making the decomposition
adversary-relative, by separating re-derivation from an \emph{irreducible} world
recall that we \emph{measure} rather than assume, and by emitting a minimal-collateral,
auditable certificate; we thus neither claim the parametric/memory split as novel nor
edit weights. Forgetting by editing the retrieval store rather than the model also links
us to RAG-unlearning \citep{ragunlearn}.

\paragraph{Deletion under dependencies.} Our formal grounding is the database line
on erasure when records are inter-derivable: P2E2's meaningful-erasure criterion
\citep{p2e2} and inference-aware deletion \citep{seqdb}. We port P2E2's
\emph{dependency-bounded} criterion to agent memory and deliberately
\emph{strengthen} it (its baseline keeps pre-existing operands present, which is
vacuous for the right to be forgotten) while inheriting its NP-hardness. This
bridge, from relational deletion theory to LLM agent memory, is one the
agent-memory works above do not make, and it is where the next section begins.

\section{Problem Formulation}\label{sec:formulation}

\paragraph{Setup.} Let $f$ be a target fact with value $v$, and let $\Sigma(v)$ be
its set of ground-truth surface forms, the ways $v$ can be written. A memory store
$S$ is a set of records, each carrying text; write $\mathrm{str}(r)$ for the set of
surface strings that appear in record $r$. Deleting from $S$ yields a surviving
store $S'\subseteq S$. An adversary $\mathcal{A}$ has full read access to $S'$,
black-box access to a base model $M$, and general world knowledge $W$; a query
$q_f$ asks for $v$. We write $\mathrm{rec}(a)=1$ when an answer $a$ reveals a
surface form in $\Sigma(v)$, as scored by the validated recovery judge
(\S\ref{sec:res-judge}). Certification is relative to a tolerance $\tau\in(0,1)$;
we use $\tau{=}0.10$ (\S\ref{sec:results}). We ask whether a deleted fact is still
recoverable, and we split recoverability into three channels that fail for
different reasons. Two of them depend on the adversary's reasoning model, so
recoverability is \emph{adversary-relative} (Def.~4).

\paragraph{Definition 1 (Residual survival).}
$\sigma_{\mathrm{res}}(f,S')=\mathds{1}\!\left[\exists\, r\in S':\;
\Sigma(v)\cap\mathrm{str}(r)\neq\varnothing\right]$: some surviving record still
physically contains a surface form of $v$. This channel does not depend on the
adversary.

\paragraph{Definition 2 (Re-derivation).}
$\sigma_{\mathrm{red}}^{\mathcal{A}}(f,S')=\mathrm{rec}\!\left(\mathcal{A}(q_f \mid
S',M,W)\right)$: the adversary reconstructs $v$ by reasoning over the surviving
store together with world knowledge, even when no record states $v$ outright.

\paragraph{Definition 3 (World recall).}
$\rho^{\mathcal{A}}(f)=\mathrm{rec}\!\left(\mathcal{A}(q_f \mid \varnothing, M, C)\right)$, where
$C\subseteq W$ is the subject's world-knowable context and the store is
\emph{empty}. Because $\rho^{\mathcal{A}}$ never reads $S$, it is
system-independent: it measures what the model recovers had $f$ never been stored
at all. We estimate it as a recovery rate over stochastic samples.

\paragraph{The three channels, in words.} A deleted home address that survives
verbatim in a leftover summary note is residual survival ($\sigma_{\mathrm{res}}$).
A deleted salary that an adversary rebuilds from a surviving job title plus a
public pay band is re-derivation ($\sigma_{\mathrm{red}}$). A deleted ``holds a
Class-3 licence,'' whose implication ``at least 18'' any model already knows on its
own, is the world recall ($\rho$). Deleting more shrinks the first two; it
cannot touch the third. Figure~\ref{fig:pipeline} shows how the probe, planner, and
certificate act on each channel.

\paragraph{Definition 4 (Recoverability, adversary-relative).}
$R_{\mathcal{A}}(f,S')=\max\!\left(\sigma_{\mathrm{res}},\,
\sigma_{\mathrm{red}}^{\mathcal{A}},\,\rho^{\mathcal{A}}\right)$. A fact is as
recoverable as its most open channel, since a rational adversary uses whichever one
is available; residual is the single channel that does not depend on $\mathcal{A}$.
The true worst case is $\sup_{\mathcal{A}}R_{\mathcal{A}}$, which we bracket with
four concrete adversaries spanning prior-generation and frontier models: $\mathcal{A}_{\text{mini}}$
(\texttt{gpt-4o-mini}), $\mathcal{A}_{4o}$ (\texttt{gpt-4o}), $\mathcal{A}_{\text{S5}}$
(Claude~Sonnet~5), and $\mathcal{A}_{\text{G5.5}}$ (GPT-5.5). A diverse panel
matters because adversary strength is not one-dimensional: the frontier models
re-derive benign facts more readily, yet they refuse sensitive attribute
inferences, so which adversary is worst varies by fact (\S\ref{sec:res-rho}).

\paragraph{Definition 5 (Floor-reaching and erasure-certification).}
A deletion is \emph{floor-reaching} for $\mathcal{A}$ when
$\sigma_{\mathrm{res}}=\sigma_{\mathrm{red}}^{\mathcal{A}}=0$: recoverability has
been driven down to the world recall, and nothing further can be deleted to
help. It is \emph{erasure-certified} when it is floor-reaching for every modeled
adversary and the world recall itself is below tolerance,
$\max_{\mathcal{A}}\rho^{\mathcal{A}}<\tau$ (the $\sup$ of Def.~4). These two come
apart in exactly one way, and that gap is the limit result of the paper.
Floor-reaching is always achievable: co-delete operands until
$\sigma_{\mathrm{red}}=0$. Certification, by contrast, fails whenever
$\rho^{\mathcal{A}}\geq\tau$, because no co-deletion can lower world recall, which never
depended on the store. A fact can then be as erased as deletion can make it and
still remain world-inferable (R11, Table~\ref{tab:certs}). Reporting the two
predicates separately is exact, and it avoids the misleading verdict ``complete but
recoverable.'' This baseline (world knowledge only, every stored operand removed)
is deliberately stronger than the dependency-bounded erasure of \citet{p2e2}, which
we discuss in \S\ref{sec:related}.

\paragraph{The optimization problem (Opt-P2E2).} Let $R_f\subseteq S$ be $f$'s own
records. We seek the smallest co-deletion that makes $f$ floor-reaching:
\begin{equation*}
\min_{D\,\subseteq\, S\setminus R_f} |D| \quad\text{s.t.}\quad
\sigma_{\mathrm{red}}^{\mathcal{A}}\!\left(f,\; S\setminus(R_f\cup D)\right) = 0 ,
\end{equation*}
where $|D|=k$ is the collateral cost; erasure-certification additionally requires
$\rho^{\mathcal{A}}<\tau$ (Def.~5). The co-deleted set $D$ is made of entailers the
user never named, which can read as a data-minimization violation, but it is the
opposite: leaving them is not minimization but failed erasure, since $f$ simply
re-derives. The planner minimizes $k$, the world recall $\rho$ bounds what no deletion can
remove, and because the certificate reports $k$, paying it (strong erasure) or not
(data minimization) becomes an explicit, audited policy choice rather than a hidden
side effect.

\paragraph{Proposition 1 (NP-hardness).} \textsc{Opt-P2E2} is NP-hard \citep{p2e2}, by a self-contained reduction
from \textsc{Minimum Hitting Set}
\citep{karp1972reducibility,garey1979computers}. Given a universe $U$ and sets
$\mathcal{S}=\{S_1,\dots,S_m\}$ with $S_j\subseteq U$, build an instance with one
fact $x_u$ per element $u\in U$ and, for each $S_j$, an entailment
$\bigwedge_{u\in S_j} x_u \vdash f$, so each $S_j$ is a re-derivation path; set
$\sigma_{\mathrm{res}}=0$ and $\rho=0$. The surviving store re-derives $f$ exactly
when some path $S_j$ survives intact, so a co-deletion $D$ achieves
$\sigma_{\mathrm{red}}=0$ exactly when $D$ meets every $S_j$, that is, when
$\{u:x_u\in D\}$ is a hitting set of $\mathcal{S}$. The minimum co-deletion
therefore equals the minimum hitting set, which is NP-hard. Vertex Cover is the
$2$-uniform special case, which is exactly our dataset's pairwise-entailment regime
(for instance, F040 re-derived from C001$\wedge$C002), so the hardness already holds
when every fact has just two operands: it is intrinsic, not an artifact of
high-arity rules. The relational hardness of \citet{p2e2} points the same way. We
therefore adopt the greedy, entailment-confidence-ordered planner of
\S\ref{sec:method}, which stays exact on the small dependency graphs seen in
practice (mean $k=1.03$, \S\ref{sec:res-planner}).

\begin{figure*}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  font=\small,
  >={Latex[length=2.4mm]},
  hdr/.style={font=\bfseries, align=center},
  cell/.style={rounded corners, draw, align=center, text width=25mm,
               minimum height=11mm, inner sep=3pt, font=\footnotesize},
  lab/.style={cell, text width=26mm},
  arr/.style={->, semithick, black!55},
  into/.style={-{Latex[length=2mm]}, semithick, black!55},
]
% ----- row bands (drawn first => behind everything) -----
\fill[teal!7]   (-0.55,-0.62) rectangle (12.15,0.62);
\fill[orange!8] (-0.55,-1.97) rectangle (12.15,-0.73);
\fill[violet!8] (-0.55,-3.32) rectangle (12.15,-2.08);
% ----- column headers -----
\node[hdr] at (4.2,1.05)  {In memory store $S$};
\node[hdr] at (7.4,1.05)  {Probe {\footnotesize(Def.\,1--3)}};
\node[hdr] at (10.6,1.05) {Minimal co-deletion};
\node[hdr] at (15.0,1.05) {Certificate {\footnotesize(Def.\,5)}};
% ----- channel labels (left column) -----
\node[lab, draw=teal!55!black,   fill=teal!16]   (LR) at (1.0,0)     {\textbf{Residual} $\sigma_{\mathrm{res}}$\\[1pt]\scriptsize surviving artifact holds $v$};
\node[lab, draw=orange!80!black, fill=orange!20] (LD) at (1.0,-1.35) {\textbf{Re-derivation} $\sigma_{\mathrm{red}}$\\[1pt]\scriptsize surviving facts entail $v$};
\node[lab, draw=violet!70!black, fill=violet!16] (LP) at (1.0,-2.70) {\textbf{World recall} $\rho$\\[1pt]\scriptsize world-knowable, \emph{irreducible}};
% ----- store column -----
\node[cell, draw=teal!55!black]   (SR) at (4.2,0)     {a duplicate / summary\\still contains $v$};
\node[cell, draw=orange!80!black] (SD) at (4.2,-1.35) {operands $o_1\!\wedge\!o_2\vdash v$\\survive};
\node[cell, draw=violet!70!black] (SP) at (4.2,-2.70) {\emph{store withheld}:\\base model $M$ + world $W$};
% ----- probe column -----
\node[cell, draw=teal!55!black]   (PR) at (7.4,0)     {exact-match scan\\$\Rightarrow\sigma_{\mathrm{res}}{=}1$};
\node[cell, draw=orange!80!black] (PD) at (7.4,-1.35) {adversary reasons\\over $S'\Rightarrow\sigma_{\mathrm{red}}{=}1$};
\node[cell, draw=violet!70!black] (PP) at (7.4,-2.70) {query $M$ alone\\$\Rightarrow\rho\in[0,1]$};
% ----- co-deletion column -----
\node[cell, draw=teal!55!black,   fill=teal!6]   (CR) at (10.6,0)     {delete the artifact\\[1pt]\textcolor{teal!45!black}{\textbf{$\sigma_{\mathrm{res}}{\to}0$}\ $\checkmark$}};
\node[cell, draw=orange!80!black, fill=orange!6] (CD) at (10.6,-1.35) {co-delete an operand\\[1pt]\textcolor{orange!60!black}{\textbf{$\sigma_{\mathrm{red}}{\to}0$}\ $\checkmark$}};
\node[cell, draw=violet!70!black, fill=violet!6, very thick] (CP) at (10.6,-2.70) {$\rho$ \emph{unchanged}:\\no deletion lowers it};
% ----- the WALL between probe and co-deletion on the world recall row -----
\fill[violet!70!black] (8.98,-3.25) rectangle (9.16,-2.15);
\node[white, font=\scriptsize\bfseries, rotate=90] at (9.07,-2.70) {WALL};
% ----- arrows: store -> probe -----
\draw[arr] (SR) -- (PR);
\draw[arr] (SD) -- (PD);
\draw[arr] (SP) -- (PP);
% ----- probe -> co-deletion (world recall row is stopped by the wall) -----
\draw[arr] (PR) -- (CR);
\draw[arr] (PD) -- (CD);
\draw[-{Bar[width=3.2mm]}, semithick, violet!70!black] (PP) -- (8.95,-2.70);
% ----- certificate (top predicates) -----
\node[cell, draw=black!55, fill=black!3, text width=42mm, minimum height=15mm] (CERT) at (15.0,-0.25)
  {$R=\max(\sigma_{\mathrm{res}},\sigma_{\mathrm{red}},\rho){=}\rho$\\[2pt]
   \texttt{floor\_reaching}: $\checkmark$ always\\
   \texttt{certified} iff $\rho<\tau$};
% ----- arrows from the two closed channels into the certificate -----
\draw[into] (CR.east) to[out=0,in=170] (CERT.west);
\draw[into] (CD.east) to[out=0,in=192] (CERT.west);
% ----- CERT reads rho against tau (down into the gauge) -----
\draw[into] (CERT.south) -- (15.0,-2.45) node[midway,right,font=\scriptsize]{$\rho$ vs $\tau$};
% ----- the world recall feeds the gauge -----
\draw[into] (CP.east) -- (12.95,-2.70);
% ----- gauge: the rho axis / tau threshold -----
\fill[teal!16]   (13.10,-2.84) rectangle (13.48,-2.56);
\fill[violet!14] (13.48,-2.84) rectangle (16.90,-2.56);
\draw[semithick] (13.10,-2.70) -- (16.90,-2.70);
\draw[thick] (13.48,-2.58) -- (13.48,-2.82);
\node[font=\scriptsize] at (13.10,-2.98) {$0$};
\node[font=\scriptsize] at (13.48,-2.98) {$\tau$};
\node[font=\scriptsize] at (16.90,-2.98) {$1$};
\fill[teal!55!black]   (13.10,-2.70) circle (1.7pt);
\fill[violet!70!black] (16.90,-2.70) circle (1.7pt);
\node[font=\scriptsize, text=teal!45!black,   align=center, anchor=south] at (13.62,-2.46) {F040 $\rho{=}0$\\certified $\checkmark$};
\node[font=\scriptsize, text=violet!60!black, align=center, anchor=south] at (16.15,-2.46) {R11 $\rho{=}1$\\uncertifiable};
\end{tikzpicture}%
}
\caption{\textbf{The \textsc{Salman} pipeline.} Recoverability decomposes into three
channels (rows), each measured by a dedicated probe: \emph{residual}
$\sigma_{\mathrm{res}}$, \emph{re-derivation} $\sigma_{\mathrm{red}}$, and the
\emph{world recall} $\rho$; overall recoverability is the worst open channel,
$R{=}\max(\sigma_{\mathrm{res}},\sigma_{\mathrm{red}},\rho)$ (Def.~4). The greedy
planner closes the two deletable channels, but $\rho$ is a \emph{wall} no co-deletion
can lower. The certificate (Def.~5) is thus \texttt{floor\_reaching} (always
achievable) yet \texttt{completeness\_certified} only when $\rho<\tau$. The gauge
shows this limit result: F040 ($\rho{=}0$) certifies, while R11 (``Class-3 licence
$\Rightarrow$ at least 18'', $\rho{=}1$) does not, so $\tau$ is auditor-chosen.}
\label{fig:pipeline}
\end{figure*}

\section{Method}\label{sec:method}

Our pipeline instantiates each object of \S\ref{sec:formulation} as a concrete,
reusable component (Figure~\ref{fig:pipeline}) behind a single
\texttt{MemorySystemAdapter} interface, so the
identical probe/planner/certificate stack runs unchanged on all three systems.

\paragraph{Probe battery.} Three probes measure the three channels of Def.~4.
The \emph{exact-match} probe scans every surviving record for a surface form of
$v$ ($\sigma_{\mathrm{res}}$). The \emph{re-derivation} probe gives the adversary
the full surviving store as notes (plus today's date and general world knowledge)
and elicits $v$ in a value-first format ($\sigma_{\mathrm{red}}^{\mathcal{A}}$).
The \emph{world-recall} probe withholds the store entirely, conditioning only on the
subject's world-knowable context, and estimates $\rho^{\mathcal{A}}$ as a recovery
rate over $n$ stochastic samples. The re-derivation \emph{probe} reads the full
surviving store (the threat model) and is distinct from the operands-only
\emph{injection control} of \S\ref{sec:res-rederiv}, which instead never stores the
target value so as to isolate $\sigma_{\mathrm{red}}$ at $\sigma_{\mathrm{res}}=0$.
Recovery is decided by the validated judge
(\S\ref{sec:res-judge}): a substring match on the ground-truth surface forms,
falling back to an LLM adjudication for numeric/semantic approximation. Three
auxiliary probes support specific results: a \emph{paraphrase} probe (residual by
implication), a retrieval-score \emph{membership-inference} probe (\S\ref{sec:limitations}),
and a \emph{KG-summary} probe for the graph system (\S\ref{sec:res-converge}).

\paragraph{Entailment detector.} Candidate re-derivation edges are scored by an
LLM-as-judge that returns, for a surviving fact (or set), the confidence that it
entails $v$ given world knowledge (\textsc{partial}/\textsc{no} are capped so a
near-miss cannot read as full entailment). The exact planner co-deletes over the
\emph{known} entailment DAG, so it does not depend on the judge to rediscover a
multi-hop path; the judge instead orders the greedy comparators and, in a deployed
store, would be the means of discovering the graph (\S\ref{sec:limitations}). Our
four-model validation (\S\ref{sec:res-judge}) motivates using Sonnet~5 for
the greedy comparator: every model has a \emph{zero} miss-rate on true multi-hop
entailers, but the weaker judges false-fire on partial operands, which would inflate
a judge-driven planner's collateral.

\paragraph{Co-deletion planner.} Against a \emph{naive} baseline that deletes only
the single surfaced record (no co-deletion; \S\ref{sec:res-residual}), our method
computes an \emph{exact} minimal co-deletion. Each target carries a boolean
\emph{entailment DAG} whose leaves are its stored operands and whose formula (built
from \textsc{and}/\textsc{or}/threshold nodes) is true exactly when the target is
still re-derivable; the smallest $D$ that closes the channel is therefore a
\emph{minimum hitting set} of that formula, i.e.\ the \textsc{Opt-P2E2} optimum of
\S\ref{sec:formulation}. The dependency graphs are small, so we solve it exactly, and
because co-deletion follows the known structure it recurses to the stored roots and
\emph{cannot miss a multi-hop entailer} for lack of an LLM-judged edge. Two
confidence-ordered greedy comparators bound the method: \emph{threshold} probes,
purges the target's own artifacts, then co-deletes entailers in descending
entailment confidence, re-probing after each until
$\sigma_{\mathrm{red}}^{\mathcal{A}}<\tau$; \emph{depth-first} co-deletes every
entailer above $\tau$ in one shot with no re-probe. The reported collateral $k=|D|$
is the measured price of the strong erasure notion (\S\ref{sec:formulation}): the
exact planner attains $k=1.03$ at an optimality gap $\le0$ (provably minimal), threshold
nearly matches it ($k=1.14$), while depth-first over-deletes roughly sixfold
($k=6.60$) (\S\ref{sec:res-planner}).

\paragraph{Certificate emitter.} For each target the emitter records the three
channel scores, the final recoverability $R_{\mathcal{A}}=\max(\cdot)$, the
heuristic, the co-deleted set and its size $k$, and the threat model. It exposes
\emph{two} machine-readable predicates from Def.~5: \texttt{floor\_reaching}
($\sigma_{\mathrm{res}}=\sigma_{\mathrm{red}}<\tau$, the deletable channels closed)
and \texttt{completeness\_certified}. The latter certifies erasure only under the
\emph{worst modeled adversary} (the $\sup$ of Def.~4, i.e.\
$\rho=\max_{\mathcal{A}}\rho^{\mathcal{A}}<\tau$), so a fact recoverable by the
stronger reasoner is reported uncertified even when the weaker one fails; this
conservative choice is why $86/250$, more than any single reasoner alone flags, are uncertifiable
(\S\ref{sec:res-rho}). A \texttt{floor\_reaching} but uncertified verdict is
precisely the limit result (Def.~5), now an audited field. The certificate thus
surfaces both the price ($k$) and the limit ($\rho$) of erasure; a recall caveat
(\S\ref{sec:res-judge}) qualifies certified verdicts, while uncertified ones are
sound lower bounds.

%========================================================================
\section{Experiments}\label{sec:results}

We evaluate on three deployed agent-memory systems spanning three architecture
families: \textbf{Mem0} \citep{mem0} (\texttt{mem0ai} 2.0.7, a deduplicating
extraction pipeline over local Chroma), \textbf{Zep/Graphiti}
\citep{rasmussen2025zep} (\texttt{graphiti-core} 0.29.2, a bi-temporal knowledge
graph over Neo4j~5.26), and \textbf{Letta/MemGPT} \citep{packer2024memgpt}
(\texttt{letta}~0.16.8, an LLM-paging agent over Postgres + pgvector). Every
component runs locally with pinned versions and disabled telemetry. The deletion
adversary has full read access to the surviving store and black-box (no
training-time) access to the base model. Wherever recovery is model-dependent it is measured against
a four-reasoner adversary panel spanning prior-generation and frontier models: %
\texttt{gpt-4o-mini}, \texttt{gpt-4o}, Claude~Sonnet~5, and GPT-5.5. The
certificate takes the worst of them (Def.~4). Because this panel already includes the
frontier GPT-5.5 and Sonnet~5, and both frontier models reproduce our recovery-judge
gold labels (\S\ref{sec:res-judge}), the results below are not an artifact of older
models. Embeddings are local \texttt{all-MiniLM-L6-v2}. The completeness threshold is $\tau=0.10$. Datasets are
synthetic subjects placed in realistic statistical contexts: 253 isolated PII facts,
298 multi-hop facts (including harder join, chain, disjunctive, diamond, and
threshold topologies whose targets are entailed only through \emph{unstored}
intermediate facts) with their stored operands and bystanders, and 250
world-recall-gradient facts across three difficulty tiers.
Recovery is scored by a validated LLM judge \citep{zheng2023judge} with a \emph{zero}
false-accept rate on the gold negatives (\S\ref{sec:res-judge}: $0/194$ for the
Sonnet~5 production judge, $95\%$ Wilson upper bound $\approx\!0.019$), so \emph{reported recovery rates are conservative lower bounds
to within that small margin}. Table~\ref{tab:master} summarizes
the headline numbers; results are presented in decreasing order of confidence.

\begin{table*}[t]
\centering
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Result} & \textbf{System} & \textbf{Headline} & \textbf{Backing control} \\
\midrule
Planner (\S\ref{sec:res-planner})        & Mem0     & 100\% complete, 0 spurious, exact $k{=}1.03$ & min-hitting-set over the entailment DAG; gap$\,{\le}\,0$ \\
world recall (\S\ref{sec:res-rho})       & base model$^\dagger$ & 86/250 \emph{uncertifiable} at residual $0$ & worst-adversary $\rho$; refusal audit \\
Re-derivation (\S\ref{sec:res-rederiv})  & Mem0, Letta & bin1 100\%, bin2 62--69\%$\to$0\% & operands-only; verified-faithful co-delete \\
Convergence (\S\ref{sec:res-converge})   & all three & 83\% / 100\% (Graphiti/Letta) & clean-delete verified first; explicit baseline \\
Residual (\S\ref{sec:res-residual})      & Mem0     & 97.2\%; 97.2\%$\to$0\% & $2{\times}2$ embedder$\times$cadence factorial \\
Judge (\S\ref{sec:res-judge})            & ---      & 0 false-accept (Sonnet~5); 0 multi-hop miss (all 4) & 351 gold + 1{,}370 pairs; 4-model, Sonnet~5 judge + pinned anchor \\
\bottomrule
\end{tabular}
\caption{Headline results, ordered by confidence; each ``backing control'' is the
check that rules out the obvious confound. $^\dagger$$\rho$ is measured against the
base model with the store withheld, so it is system-independent, which is the limit
result; those certificates were emitted during the Mem0 run.}
\label{tab:master}
\end{table*}

\subsection{Minimal Co-Deletion: The Planner}\label{sec:res-planner}

On all 298 multi-hop targets, the planner reads each target's entailment DAG and deletes a
\emph{minimum hitting set} of it: the fewest surviving operands whose removal makes the
target un-re-derivable. This closes the re-derivation channel on \emph{every} target
(\textbf{100\% completeness}) while touching almost nothing extra: \textbf{zero spurious
bystander deletions} and a mean of only \textbf{$k=1.03$} operands deleted per target.
Because that hitting set is computed from the known structure, $k$ is not merely small but
\emph{provably minimal}---the measured optimality gap is $\le0$ on every topology (aggregate $-0.067$). It sits above one
only because the harder topologies demand it: an ``at least two of three'' target cannot be
closed by deleting a single operand, so its optimum is $k^\star{=}2$, and the planner pays
exactly that.

\textbf{Why ``zero spurious'' is the real test.} Each target's candidate pool mixes its
true operands with fixed bystanders the planner must never delete. The exact planner
deletes only operands named in the target's DAG, so selectivity is structural; \textbf{zero
spurious across all 298 targets} confirms the recovered structure never mistakes a bystander
for an operand.

\textbf{What structure buys, measured.} Two greedy comparators bound the exact planner
(\S\ref{sec:method}). \emph{Threshold}---a confidence-ordered greedy that re-probes and
stops early---nearly matches the optimum ($k=1.14$, gap $\approx0$) but leans on the
entailment judge's selectivity, which is why it uses Sonnet~5, the judge that
false-fires least (\S\ref{sec:res-judge}); weaker judges such as \texttt{gpt-4o-mini}
mistake partial operands for sufficient ones and would inflate $k$. \emph{Depth-first}---a single probe-free pass that
sweeps in every above-$\tau$ candidate---also reaches full completeness but at roughly
\textbf{six times} the collateral ($k=6.60$) and with \textbf{over a thousand spurious
deletions}, because without re-probing it cannot tell when to stop. The exact-to-depth-first
gap is the measured price of ignoring structure; the near-zero exact-to-threshold gap shows
a good greedy recovers most of it on these small graphs. The one place the greedy visibly
pays is the disjunctive topology $((A\vee B)\wedge C)$, where it removes both OR-branches
($k{=}1.63$) when deleting the single shared operand ($k^\star{=}1$) would suffice---a
concrete case where structure-awareness beats confidence ordering. The lower baseline is
naive single-record deletion (\S\ref{sec:res-residual}), which removes one record and closes
no re-derivation path.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcc@{}}
\toprule
field & \textsc{complete} & \textsc{incomplete} \\
\midrule
example (system)                 & F040 (Letta)   & R11 (Mem0 run) \\
\texttt{residual\_survival}      & 0.0            & 0.0 \\
\texttt{rederiv\_with\_operands} (pre)  & 1.0     & --- \\
\texttt{re\_derivation\_score} (final)  & 0.0     & 0.0 \\
\texttt{parametric\_risk\_rho}   & 0.0            & \textbf{1.0} \\
\texttt{facts\_co\_deleted} ($k$)& C001,C002 (2)  & --- (0) \\
\texttt{final\_recoverability}   & 0.0            & 1.0 \\
\texttt{floor\_reaching}         & true           & \textbf{true} \\
\texttt{completeness\_certified} & \textbf{true}  & \textbf{false} \\
\bottomrule
\end{tabular}%
}
\caption{Two real certificates. \emph{F040} (\textsc{complete}, Letta): re-derivation
was open before co-deletion (\texttt{rederiv\_with\_operands}$=1.0$) and closed after
(\texttt{re\_derivation\_score}$=0.0$). \emph{R11} (\textsc{incomplete}, Mem0):
residual and re-derivation are both $0$, yet the base-model world recall $\rho{=}1.0$ forces
\textsc{incomplete}, the limit result. Both are \texttt{floor\_reaching}; only F040
is \texttt{completeness\_certified}. Different systems, one format.}
\label{tab:certs}
\end{table}

\subsection{The World Recall and the Limit of Certifiable Deletion}\label{sec:res-rho}

The world recall $\rho$ is the one channel deletion cannot touch: the rate at which
the base model reconstructs a value from world-knowable context alone, with the store
withheld. We \emph{measure} it rather than assume it, on 250 facts written across a
low/mid/high difficulty gradient. We always report the \emph{measured} world recall, not the
authored tier; where the two disagree we keep the fact and report the measurement, since
the authored tier is only a hypothesis and $\rho$ is the evidence.

We score each fact against the \emph{worst} of our four adversaries. This sorts the 250
facts into three groups (Table~\ref{tab:rho}): most are certifiable, a middle
band sits in an intermediate range, and the rest hit a hard tier, with the mean world recall
rising along the difficulty gradient. The headline follows: \textbf{86 of 250 facts are
not erasure-certifiable} at $\tau{=}0.1$, \emph{even though their residual survival is
zero and every re-derivation path was closed}. This is the limit result: once a value is
inferable from world knowledge no one can delete, no co-deletion can certify its erasure.
And because the world recall is a smooth gradient rather than an all-or-nothing split, the
uncertifiable count moves with $\tau$, which makes $\tau$ auditor-chosen:
setting it is choosing how much residual world-inferability to tolerate.

\textbf{Which adversary is worst varies by fact.} The four-model panel is not a
formality. On \emph{benign} facts the frontier models re-derive more readily, so they set
the worst case. But on \emph{sensitive} high-tier facts, such as a national-service
status that implies sex or a licence class that implies a minimum age, the frontier
models often \emph{refuse} the inference outright, which turns their measured $\rho$ into
a lower bound and leaves the older \texttt{gpt-4o} as the worst adversary there. Stronger
models are thus not uniformly stronger adversaries, and taking the worst over a diverse
panel is what captures that. This positive world recall does not contradict the $\rho{=}0$ we
report for the multi-hop facts: there the world-knowable part is inert without the
fictional stored operand and collapses under co-deletion, whereas here the context is
itself world-knowable and pins the target down, so its recovery sits in the irreducible
world recall rather than the deletable channel.

\textbf{Control.} By construction the $\rho$ probe sees \emph{only} the subject's
world-knowable context and never the stored facts, so co-deletion provably cannot lower
it. Three checks harden the gradient: a numeric-tolerance sweep shows the world recall survives
at every tolerance; a base-rate versus context-lift split separates real inference from
mere prior plausibility; and a refusal audit confirms that $\rho$ on sensitive
attributes is an adversarial lower bound. Two caveats bound the reading. Because $\rho$
is a rate over a handful of stochastic samples per fact, the count is a Monte-Carlo
estimate good to a fact or two; and a \textsc{complete} certificate is complete only
modulo the recovery judge's imperfect recall, while \textsc{incomplete} and aggregate
figures stay safe lower bounds. Per-tier means, refusal counts, and sample sizes are in
the supplementary material.

\begin{table}[t]
\centering
\small
\begin{tabular}{@{}llcc@{}}
\toprule
worst-adversary $\rho_{\max}$ & meaning & \#facts & certifiable? \\
\midrule
$\le\tau$ ($\tau{=}0.1$) & certifiable & 164 & \checkmark \\
$\tau$ to $0.5$          & intermediate band & 41 & no \\
$\ge 0.5$                & hard tier        & 45 & no \\
\midrule
\multicolumn{2}{@{}l}{total} & 250 & 164 \\
\bottomrule
\end{tabular}
\caption{World recall $\rho$ on $n{=}250$ facts ($8$ samples each), binned by the
worst of the four adversaries, $\rho_{\max}$. The world recall is \emph{graded}: a $41$-fact
intermediate band sits between the certifiable $164$ and the hard tier $45$, so the
$86/250$ uncertifiable count moves with $\tau$ (auditor-chosen). Measured band disagrees
with authored tier for $77/250$ facts.}
\label{tab:rho}
\end{table}

\subsection{Decomposing Recovery: Re-Derivation at Zero Residual}\label{sec:res-rederiv}

To isolate re-derivation from residual survival, we use an \emph{operands-only control}:
we inject a target's entailing facts but never store the target value itself, so residual
survival is zero by construction. Whatever the adversary now recovers, it recovered by
\emph{reasoning}, not by reading a leftover copy. We sort the targets into two bins by
what that reasoning needs: \emph{stored-alone}, where the surviving operands suffice, and
\emph{stored+world}, where they must be combined with world knowledge.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lccccc@{}}
\toprule
re-derivation (\%) & mini & 4o & S5 & G5.5 & post \\
\midrule
Mem0 bin~1 (stored-alone)       & 100 & 97 & 100 & 100 & 0 \\
Mem0 bin~2 (stored+world)       & 62  & 69 & 69  & 66  & 0 \\
Mem0 structured (5 topologies)  & 100 & 100 & 100 & 100 & 0 \\
Letta bin~2 (stored+world)      & 62  & 65 & 65  & 68  & 0 \\
\bottomrule
\end{tabular}
\caption{Re-derivation rate at residual $0$ (operands-only control), by adversary
(\texttt{gpt-4o-mini}, \texttt{gpt-4o}, Sonnet~5, GPT-5.5); ``post'' is after
co-deleting the operands. The five harder topologies (join, chain, disjunctive
$((A\vee B)\wedge C)$, diamond, and at-least-$k$-of-$n$ threshold) re-derive at $100\%$
and also collapse to $0$; the world recall is $\rho=0$ throughout. Extended detail
(confidence intervals, Letta faithfulness) is in the supplementary material.}
\label{tab:rederiv}
\end{table}

On Mem0, across all four adversaries, the target is re-derived on nearly every
stored-alone fact and on about two-thirds of the stored+world facts, and co-deleting the
operands drives both to zero, down to a world recall of $\rho=0$
(Table~\ref{tab:rederiv}). Adversary strength is not one-dimensional: the worst
stored+world rate the certificate must anticipate is set by \texttt{gpt-4o} and Sonnet~5
(both $69\%$), while the newest GPT-5.5 sits lower ($66\%$)---a larger model is not
uniformly a stronger re-deriver, which is exactly why we take the worst over a diverse panel. The
five harder topologies stress the planner's reach: in each, the target is entailed not by
stored facts but by \emph{derived} intermediates that are themselves entailed by stored
roots, and for the disjunctive and threshold targets by more than one operand set. Every
adversary re-derives all of them, and co-deletion again drives recovery to zero---but only
because the planner recurses \emph{past} the unstored direct entailers to the stored roots
and deletes a true minimum hitting set, both of which a one-hop or delete-everything
planner would get wrong.

The whole control ports to Letta with the same pattern and the same collapse to zero after
co-deletion. The probe is not firing indiscriminately: on a fact like F043 (a
Catholic-High-School-to-Bishan geographic leap) the two backbone reasoners never make the
inference, and only the frontier models do---an honest per-fact negative for the weaker
adversaries, and another reminder that which adversary is worst varies by fact. Re-derivation is thus a recovery channel orthogonal to
residual survival, and minimal co-deletion closes it down to the world recall.

\textbf{Control.} We bin recovery by mechanism and never average across bins, and run all
four reasoners because re-derivation is model-dependent. On Letta this matters twice over:
each co-deletion goes through the direct, deterministic \texttt{passages.delete} by
captured passage id, not a conversational request, and we verify every operand passage is
actually gone. So ``re-derivation closed'' cannot be confused with ``the delete silently
did not run,'' which keeps this result cleanly separate from the agent-mediated-deletion
finding of \S\ref{sec:res-converge}. Confidence intervals and the full multi-level and
Letta-faithfulness breakdowns are in the supplementary material.

\subsection{Generalization: Three Families, One Phenomenon}\label{sec:res-converge}

\begin{figure*}[t]
\centering
\begin{tikzpicture}[
  font=\small,
  surf/.style={rectangle, rounded corners, draw, minimum width=24mm, minimum height=10mm, align=center},
  cleared/.style={surf, fill=black!4},
  residual/.style={surf, draw=red!70!black, fill=red!10, very thick},
  agent/.style={rectangle, rounded corners, draw, fill=blue!5, minimum width=36mm, minimum height=9mm, align=center},
  instr/.style={align=center, text width=48mm},
  arr/.style={-{Latex[length=2.2mm]}, thick},
  faith/.style={align=center}
]
% ----- Panel A: explicit -----
\node[instr] (iA) {\textbf{Explicit instruction}\\``delete X from core \textbf{and} archival''};
\node[agent, below=3mm of iA] (aA) {agent $\to$ \texttt{memory\_replace}\\\footnotesize(acts on both surfaces)};
\node[cleared, below=6mm of aA, xshift=-15mm] (cA) {\textbf{core block}\\\footnotesize(X removed)};
\node[cleared, below=6mm of aA, xshift=15mm]  (rA) {\textbf{archival}\\\footnotesize(X removed)};
\draw[arr] (aA) -- (cA);
\draw[arr] (aA) -- (rA);
\node[faith, below=18mm of aA] {\textbf{100\% faithful}\\\footnotesize both surfaces cleared};
% ----- Panel B: vague -----
\node[instr, right=30mm of iA] (iB) {\textbf{Vague, realistic request}\\``I'm not comfortable with you keeping X''};
\node[agent, below=3mm of iB] (aB) {agent $\to$ \texttt{memory\_replace}\\\footnotesize(acts on core only)};
\node[cleared, below=6mm of aB, xshift=-15mm]  (cB) {\textbf{core block}\\\footnotesize(X removed)};
\node[residual, below=6mm of aB, xshift=15mm]  (rB) {\textbf{archival}\\{\large\textbf{X}}~\footnotesize(retained)};
\draw[arr] (aB) -- (cB);
\draw[arr, red!70!black, dashed] (aB) -- (rB) node[midway, right, red!70!black, font=\footnotesize] {misses};
\node[faith, below=18mm of aB] {\textbf{0\% faithful} / 100\% archival residue\\\footnotesize core cleared, archival silently missed};
\end{tikzpicture}
\caption{\textbf{Same Letta agent, opposite outcomes by phrasing.} An explicit
dual-surface instruction (left) makes the agent scrub \emph{both} the core block
and the archival store: $100\%$ faithful. A vague, deployment-realistic request
(right) makes the \emph{same} agent call \texttt{memory\_replace} on the core block
it reasons over and \emph{silently miss} archival: $0\%$ faithful, $100\%$
archival residue. The gap is phrasing and surface coverage, not capability: the
explicit baseline proves the agent \emph{can} clear both. This is the agent loop
ForgetEval \citep{forgeteval} bypassed by writing archival endpoints directly.}
\label{fig:agentloop}
\end{figure*}

Residual survival is not a Mem0 quirk: it appears in all three architecture families,
each through a \emph{different by-design mechanism}. We run these two cross-system probes
on a small, high-stakes set of 30 sensitive-PII targets per system (see
\S\ref{sec:limitations}).

In \textbf{Graphiti}, the explicit \texttt{remove\_episode} is clean at the graph level,
hard-deleting the episode, its \texttt{RELATES\_TO} edges, and any orphaned entities (we
verify this first). Yet the deleted value lives on in \textbf{stale entity and community
summaries that the system never recomputes on deletion}. Edge residue is low, but
\textbf{summary residue reaches $83\%$} (we count a value present in any surviving edge or
summary, regardless of its bi-temporal validity fields, consistent with the
full-read-access threat model). In \textbf{Letta}, deletion is agent-mediated across two
surfaces, the in-context \emph{core} blocks and the \emph{archival} passages, and the
outcome depends on how the request is phrased. An explicit dual-surface instruction is
\textbf{$100\%$ faithful}. But a \emph{vague, deployment-realistic} request (``I'm not
comfortable with you keeping X''), with the fact also sitting in archival, makes the agent
scrub the core block it is reasoning over and silently miss archival: \textbf{$0\%$
faithful, $100\%$ archival residue} (Figure~\ref{fig:agentloop}). The agent even confirms a
successful forget while the fact persists, an over-compliant acknowledgment consistent with
known sycophancy \citep{sharma2024sycophancy}. Agent-mediated deletion is thus
\emph{surface-incomplete} and instruction-dependent.

\textbf{Novelty and corroboration.} This agent-loop failure is genuinely new:
ForgetEval \citep{forgeteval} deliberately bypassed Letta's agent loop
(addressing the archival REST endpoints directly, keeping the LLM out of the
recall hot path), so we test exactly the path it excluded. Conversely,
ForgetEval's own per-system notes \emph{independently corroborate} our convergence
thesis: its observation that Graphiti ``sheds surface forms / synthesised fact
strings'' is our stale-summary residue, and that Mem0's router prefers
``link-and-keep over delete-old'' is our duplication residue.

\textbf{Control.} For Graphiti, verifying the clean hard delete \emph{first}
establishes that the residue is stale summaries, not failed edge removal. For
Letta, the $100\%$-faithful explicit baseline isolates the failure to
phrasing/surface coverage, not to a missing capability.

\subsection{Residual Survival and the Duplication Mechanism}\label{sec:res-residual}

Naive single-record deletion on Mem0 is residually incomplete for one reason: the store
\emph{silently duplicates} facts, so deleting the single surfaced record leaves copies
behind. That single-record delete leaves the value present in \textbf{$97.2\%$} of
isolated PII targets, whereas an \emph{artifact-aware} delete, one that purges every
record carrying the value, drives residual survival \textbf{to $0\%$}. Both arms run on
the same isolated PII set, whose probe values are all globally distinct so that one
fact's measurement cannot contaminate another's, and the two runs land independently on
the same rate, a small replication of the effect.

\textbf{Control.} A $2{\times}2$ factorial over embedder and injection cadence shows
value duplication in \emph{all four} cells (row inflation $+174$--$182\%$), with both
byte-identical and paraphrase-variant duplicates present throughout; restricting to facts
whose value is unique in the corpus, so vocabulary overlap cannot be miscounted, the
duplication persists at a high rate ($80$--$82\%$ incidence). It is therefore a semantic-dedup \emph{design limitation}, not an artifact of
our embedder or an injection-timing race, and it is independently documented in Mem0's
own issue tracker. These rates are not in tension with ForgetEval's Mem0 numbers: their
figure is forget-command success in top-$k$ retrieval, whereas ours is post-deletion
recoverability by \emph{any} channel, complementary measures of necessary versus
sufficient conditions for completeness. Exact cell rates and issue references are in the
supplementary material.

\subsection{Judge Validation}\label{sec:res-judge}

Two LLM judges do load-bearing work, one deciding whether an answer counts as
\emph{recovery} and the other whether a set of surviving facts \emph{entails} the
target, so we validate both against hand-labeled gold.

We validate both judges on large gold sets and under \emph{all four} adversary models,
reporting rates with Wilson intervals and inter-model agreement (Cohen's $\kappa$).

The \textbf{recovery judge} is checked on $351$ labeled cases ($194$ negatives) spanning
exact, paraphrase, numeric within- and out-of-tolerance, wrong-value, opposite, and
refusal answers, including $22$ curated grey-zone cases (reconstruction phrasings,
category-vs-specific, hedges, one-token flips). The frontier \textbf{Sonnet~5} production
judge makes \emph{zero} false accepts in the $194$ gold negatives ($0\%$, $95\%$ Wilson
upper bound $\approx\!0.019$) at $0.98$ recall; its errors run the safe way---scoring a
close numeric near-miss as a non-recovery, never the reverse---so reported recovery rates
are lower bounds to within that small margin. On this larger, harder gold the pinned
models slip: \texttt{gpt-4o} makes one false accept ($0.5\%$) and \texttt{gpt-4o-mini}
four ($2.1\%$), while GPT-5.5 also reaches zero. We therefore run the most capable model,
Sonnet~5, as the production judge---empirically the safest on the gold---and keep the
pinned \texttt{gpt-4o-mini}/\texttt{gpt-4o} rates as a reproducibility anchor, since
neither Sonnet~5 nor GPT-5.5 has a dated snapshot a reviewer could pin against.

The \textbf{entailment judge} is checked on $1{,}370$ pairs: positives (full operand
sets, single- and multi-hop) and adversarial near-misses (single operands that do not
suffice). Two findings settle its use. First, \emph{every} model has a \textbf{zero
miss-rate on true multi-hop entailers}---none silently drops a genuine multi-hop
dependency, exactly the error a co-deletion planner most needs to avoid, and the exact
planner removes even that risk structurally (\S\ref{sec:method}). Second, on the
near-misses the models diverge sharply: \texttt{gpt-4o-mini} treats $75.7\%$ of
insufficient single operands as sufficient, \texttt{gpt-4o} $45.5\%$, GPT-5.5 $30.5\%$,
and Sonnet~5 only $3.4\%$. A high rate here would inflate a \emph{judge-driven} planner's
collateral, so the greedy comparator uses Sonnet~5, the judge that false-fires least; the
exact planner, co-deleting over the known DAG, is unaffected. Sonnet~5 is thus both a
panel adversary and the production judge for recovery and entailment---an overlap we
accept because the judge's guarantee is gold-validation ($0\%$ false-accept on the
$351$-case ground-truth-by-construction gold), not the independence of its calls, since
same-model errors correlate across calls in any case. The pinned \texttt{gpt-4o} numbers
are retained alongside as a reproducibility anchor. Full per-model rates and $\kappa$ are
in the supplementary material.

%========================================================================
\section{Limitations}\label{sec:limitations}

We state the boundaries of the evidence, leading with the issue we caught and fixed.

\paragraph{Membership inference: retracted, then powered.} Our initial
membership-inference probe was underpowered (only $n{=}6$ members), so we withdrew that
reading and re-ran it properly on $n{=}253$ members, each with three matched near-twins. The powered picture is nuanced. The test clearly has
power: before any deletion, an adversary tells stored facts from unstored ones well
above chance (AUC $0.66$). Naive single-record deletion barely dents that signal
(AUC $0.66$). Artifact-aware deletion pushes it down close to chance (AUC $0.51$), but
not conclusively: the confidence interval just includes $0.5$, while a permutation test
is marginally significant ($p{=}0.04$). We therefore make only the defensible claim, that
artifact-aware deletion sharply reduces the membership signal but, at this sample size,
we cannot show it is fully erased.

\paragraph{Our numbers inherit the judges' limits.} Both LLM judges are validated on
large gold sets across four models (\S\ref{sec:res-judge}) but imperfect. On single
insufficient operands the entailment judge can mistake a partial operand for a sufficient
one, which is why the greedy comparator uses the least-false-firing judge, Sonnet~5, and
the method itself co-deletes over the known DAG rather than trusting the judge. The
recovery judge (Sonnet~5) makes \emph{zero} false accepts on the $194$ gold negatives
($95\%$ Wilson upper bound $\approx\!0.019$), so the ``lower bound'' framing holds to
within that small margin; every model, though, has a zero miss-rate on true multi-hop
entailers.

\paragraph{The world recall is a lower bound where it matters most.} We estimate the
world recall by asking the base model for the value, and for sensitive attributes
the model often refuses. A refusal makes the measured $\rho$ an underestimate of true
world-inferability, precisely for the attributes a privacy audit cares about most. A
more willing adversary could only push the uncertifiable count higher, never lower, so
the limit result is conservative.

\paragraph{The planner needs a known entailment graph.} Our decomposition assumes the
entailment graph is given, as it is in our controlled setup, where we injected the
operands ourselves. A deployed store hands the auditor no such graph and it is hard to
recover, so the planner can only close paths through entailers it can see. An unmodeled
entailer stays open, and recovering the full graph of a live memory is its own open
problem. Even at zero residual and re-derivation, then, completeness is relative to what
the auditor knows.

\paragraph{The cross-system evidence is small.} The convergence result rests on $n{=}30$
high-stakes targets per system. The three mechanisms are qualitatively distinct and each
is independently corroborated \citep{forgeteval}, but the per-system rates are
indicative rather than precise at this scale. All subjects are synthetic, so recoveries
reflect re-derivability, not memorization of real people.

\section{Conclusion and Future Work}\label{sec:conclusion}

Deleting a record is not erasing a fact. We gave a causal decomposition of
post-deletion recoverability, a planner that closes the deletable channels with
minimal collateral, and a certificate that states, per fact, both the cost of
erasure and whether it can be certified at all. Across three memory architectures
the same gap recurs through different by-design mechanisms, and a \emph{measured}
world recall sets a hard limit on what any deletion can promise.

That limit is the paper's broadest implication. When a fact is recoverable from
world knowledge alone ($\rho\geq\tau$), as $86$ of our $250$ gradient facts are under
the strongest modeled adversary, no external-store deletion can erase it: the value
is reconstructed without ever reading the store. Right-to-be-forgotten compliance
then cannot be a deletion operation; it requires not storing the fact in the first
place, or accepting and disclosing the residual world recall. Deletion engineering closes
the re-derivation and residual channels, not the world-recall one.

One extension is immediate: a fourth memory architecture, to test whether the
three-family convergence is universal. The deeper open problem is certifying erasure
against an unbounded adversary rather than a modeled bracket: the world recall we measure is
a lower bound, so the true limit of forgetting in agent memory may be higher still.

\paragraph{Broader impact.} Our goal is to make deletion in agent memory
\emph{auditable}: the certificate tells an operator when a fact is truly gone and, more
usefully, when it cannot be. The recovery probes that expose incompleteness are the
same techniques an adversary would use, but they run on synthetic subjects only, so no
real personal data is at risk; we intend them as a tool for builders to verify erasure,
not to attack deployed stores. The limit result also carries a design message: when a
value is world-inferable, privacy is better served by not storing it than by deleting
it after the fact.

%========================================================================
% NOTE: do NOT add \bibliographystyle here. aaai2027.sty already issues
% \bibliographystyle{aaai2027} internally; a second one triggers the bibtex error
% "Illegal, another \bibstyle command". Just \bibliography{...} below. See AAAI_SETUP.md.
\bibliography{references}

\end{document}
